---
title: "transform data to update the URL"
author: "John Little"
date: "`r Sys.Date()`"
output: html_notebook
---

The PowerHouse Museum updated their target website.  Thus their "Persistent Links" are no longer accurate for web scraping.  They do auto-forward to the new target URL.  But, for purposes of the tutorial, I'm updating "Persistent link to the latest URL"

# Load Packages

```{r}
library(tidyverse)
```

# Load Data

```{r}
temp <- read_delim("data/bicycle-subset-phm-collection.tsv", 
    "\t", escape_double = FALSE, trim_ws = TRUE)

temp
```

# Transform Data

update the link here by creating a new variable, "new_link"

```{r}
new_bicycle_metadata <- temp %>% 
  mutate(new_link = paste0(
    "https://collection.maas.museum/object/",
    str_split(`Persistent Link`, "=", 
              simplify = TRUE)[,2]))

new_bicycle_metadata <- new_bicycle_metadata %>% 
  select(1:10, 17, everything())
```

# Write data as TSV

```{r}
write_tsv(new_bicycle_metadata, "data/new_url.tsv")
```


# Document Session Info

```{r}
devtools::session_info()
```

