---
title: "Link Documentation to blogdown Static Directory"
---

**README**: This documentation should only have to be done once.  It creates a symbolic link from the Rfun website's static directory (blogdown's static directory) to the HTML documentation generated here by using RStudio's "Build website feature".



```{r}
library(fs)
library(here)
library(glue)
```

```{r}
dir_ls()
```

## `pwd()`
```{r}
#docstoo <- dir_create("docs2")
#dir_copy("docs", here())

here()
```

## Source dir

```{r}
source_dir <- "_book"
```

## Target dir

```{r}
#file_copy("docs", "docs2")
#dir_copy("docs", "docs2")
#dir_copy("docs", here("docs2"))
#dir_delete("docs2")
#here()
#dir_copy(here("docs"), here("docs2"))

up_1_dir <- ".."
#dir_workDesktop_funky1 <- "R"
#dir_workDesktop_funky2 <- "github"
dir_blogdown_target <- "libjohn.github.io" 
#dir_blogdown_static_target <- "static"
dir_map_docs_target <- "openrefine"

target_dircopy_location <- glue(up_1_dir,
                                #dir_workDesktop_funky1,
                                #dir_workDesktop_funky2,
                                dir_blogdown_target,
                                #dir_blogdown_static_target,
                                dir_map_docs_target,
                                .sep = "/")

#target_dircopy_location
here(target_dircopy_location)

#dir_copy(here("docs"), here(target_dircopy_location))
#file_move(here(target_dircopy_location, "docs"), 
#          here(target_dircopy_location, "map"))
```

## Create Link

This next chunk will create a symbolic link from the local dir to target location (`target_dircopy_location`)

## Make Link / Procedure

To run, **and to avoid** unnecessary **messiness**, append `{r}` to the code chunk below.  Then run the chunk.  

But, **only need to run this once!**   Because, this creates a symbolic link from the source to the target.


```
link_create(here(source_dir),
            here(target_dircopy_location))
```

